\mpx{}: Mixed Precision Training for JAX

### 2025-05-20 doc/paper/main.tex
added pypi configs

\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{enumitem}

\title{MPX: Mixed Precision Training for JAX}
\author{}
\date{}

@@

\section{Introduction}
This paper presents \textbf{mpx}, a tool for training JAX models using mixed precision. The library extends the capabilities of JMP (JAX Mixed Precision) \cite{jmp}, addressing its limitations in handling arbitrary PyTrees and compatibility with models developed using Equinox \cite{equinox}. By leveraging Equinox's flexibility, mpx provides a solution that works with any PyTree structure.

@@

\section{Implementation Details}
mpx is to provides transformations that allow users to transform their existing training pipeline into mixed precision.
For this, it provides several functions that allow to cast

The mpx library provides essential transformations for mixed precision training while maintaining JAX's low-level approach. Key components include:

\begin{enumerate}
    \item \textbf{Transformations to Cast PyTrees}: mpx features the following functions to cast arbitrary PyTrees: \texttt{cast\_tree(tree, dtype)} \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_float16(x)}, \texttt{cast\_to\_bfloat16(x)}, \texttt{cast\_to\_float32(x)}. All these functions cast all leaves of the input that are JAX arrays and of type float to the corresponding float datatype. All other leaves, including arrays that are of non-float types, like int32, remain unchanged. 
    \item \textbf{Transformations to Cast Functions}: mpx contains a transformation \texttt{cast\_function(func, dtype, return\_dtype=None)} for functions. This transformation returns a function that casts all its inputs to the desired input datatype (using \texttt{cast\_tree(tree, dtype)}), calls the function and then casts the outputs of the function. Moreover, mpx contains \texttt{force\_full\_precision(func, return\_dtype)}, which forces a function to perform its computations with full precision. This is important as some operations, such as sum, mean or softmax, are sensitive to overflows when calculated in float16.
    \item \textbf{Transformations to Cast Gradients}: mpx contains the Equinox equivalents \texttt{filter\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} and \texttt{filter\_value\_and\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} that calculate the gradient of a function using mixed precision with loss scaling (as described above). Additional to calculating the gradient, the functions also perform the automatic adaption of the loss scaling value.
    These drop-in replacements allow users to reuse their existing Equinox training pipelines without major changes to their structure (cf. Section todo).
\end{enumerate}

@@

\subsection{Optimizer}
mpx works with all optax optimizers. However, as explained above, one might need to skip optimizer updates if gradients became infinite due to loss scaling.
The  \texttt{optimizer\_update(model, optimizer, optimizer\_state, grads, grads\_finite)} function handles model updates based on gradient finiteness.
This means, instead of calling \texttt{optimizer.update}, followed by \texttt{eqx.apply\_updates} as done in regular Equinox training pipelines, one just have to call \texttt{mpx.optimizer\_update}.

\section{Example}
I t

@@

\begin{thebibliography}{9}
\bibitem{jmp} 
JMP: JAX Mixed Precision
\newblock \url{https://github.com/google-deepmind/jmp}

\bibitem{equinox}
Equinox: Neural Networks in JAX
\newblock \url{https://docs.kidger.site/equinox/}

\bibitem{nvidia_amp}
NVIDIA Automatic Mixed Precision
\newblock \url{https://developer.nvidia.com/automatic-mixed-precision}

\bibitem{mixed_precision_paper}
P. Micikevicius et al.
\newblock ``Mixed Precision Training''
\newblock arXiv:1710.03740, 2017
\end{thebibliography}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[hidelinks]{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{enumitem}

\newcommand{\mpx}{\textsc{MPX}}

\title{\mpx{}: Mixed Precision Training for JAX}
\author{}
\date{}

@@

\section{Introduction}
This paper presents \mpx{}, a tool for training JAX models using mixed precision. The library extends the capabilities of JMP (JAX Mixed Precision) \cite{jmp}, addressing its limitations in handling arbitrary PyTrees and compatibility with models developed using Equinox \cite{kidger2021equinox}. By leveraging Equinox's flexibility, \mpx{} provides a solution that works with any PyTree structure.

@@

\section{Implementation Details}
\mpx{} is to provides transformations that allow users to transform their existing training pipeline into mixed precision.
For this, it provides several functions that allow to cast

The \mpx{} library provides essential transformations for mixed precision training while maintaining JAX's low-level approach. Key components include:

\begin{enumerate}
    \item \textbf{Transformations to Cast PyTrees}: \mpx{} features the following functions to cast arbitrary PyTrees: \texttt{cast\_tree(tree, dtype)} \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_half\_precision(x)}, \texttt{cast\_to\_float16(x)}, \texttt{cast\_to\_bfloat16(x)}, \texttt{cast\_to\_float32(x)}. All these functions cast all leaves of the input that are JAX arrays and of type float to the corresponding float datatype. All other leaves, including arrays that are of non-float types, like int32, remain unchanged. 
    \item \textbf{Transformations to Cast Functions}: \mpx{} contains a transformation \texttt{cast\_function(func, dtype, return\_dtype=None)} for functions. This transformation returns a function that casts all its inputs to the desired input datatype (using \texttt{cast\_tree(tree, dtype)}), calls the function and then casts the outputs of the function. Moreover, \mpx{} contains \texttt{force\_full\_precision(func, return\_dtype)}, which forces a function to perform its computations with full precision. This is important as some operations, such as sum, mean or softmax, are sensitive to overflows when calculated in float16.
    \item \textbf{Transformations to Cast Gradients}: \mpx{} contains the Equinox equivalents \texttt{filter\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} and \texttt{filter\_value\_and\_grad(func, scaling, has\_aux=False, use\_mixed\_precision=True)} that calculate the gradient of a function using mixed precision with loss scaling (as described above). Additional to calculating the gradient, the functions also perform the automatic adaption of the loss scaling value.
    These drop-in replacements allow users to reuse their existing Equinox training pipelines without major changes to their structure (cf. Section todo).
\end{enumerate}

@@

\subsection{Optimizer}
\mpx{} works with all optax optimizers. However, as explained above, one might need to skip optimizer updates if gradients became infinite due to loss scaling.
The  \texttt{optimizer\_update(model, optimizer, optimizer\_state, grads, grads\_finite)} function handles model updates based on gradient finiteness.
This means, instead of calling \texttt{optimizer.update}, followed by \texttt{eqx.apply\_updates} as done in regular Equinox training pipelines, one just have to call \texttt{\mpx{}.optimizer\_update}.

\section{Example}
Here, we provide an example and show which parts in a training pipeline need to be changed for mixed precision training.

\section{Model Implementation}
For the largest part, the implementation of the model must not be changed. 
As \mpx{} works with arbitrary PyTrees, every Toolbox that defines their model/parameters as PyTrees, like Flax~\cite{flax2020github} or Equinox~\cite{kidger2021equinox} can be used in conjunction with \mpx{}.

@@

\bibliographystyle{plain}
\bibliography{references}

% \begin{thebibliography}{9}
% \bibitem{jmp} 
% JMP: JAX Mixed Precision
% \newblock \url{https://github.com/google-deepmind/jmp}

% \bibitem{equinox}
% Equinox: Neural Networks in JAX
% \newblock \url{https://docs.kidger.site/equinox/}

% \bibitem{nvidia_amp}
% NVIDIA Automatic Mixed Precision
% \newblock \url{https://developer.nvidia.com/automatic-mixed-precision}

% \bibitem{mixed_precision_paper}
% P. Micikevicius et al.
% \newblock ``Mixed Precision Training''
% \newblock arXiv:1710.03740, 2017
% \end{thebibliography}
